---
title: 'Part 1: Text mining and exploratory analysis'
author: "Eva Y"
date: '2018-06-06'
output: 
  html_document: 
    keep_md: yes
editor_options: 
  chunk_output_type: console
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=8, fig.height=8,
                      echo=FALSE, warning=FALSE, message=FALSE)

knitr::opts_knit$set(root.dir = "../")
```

## Goals

The purpose of this document is perform text mining and exploratory analysis on chat sessions provided by Comm100. This includes measuring word frequency, word length, lexical diversity, and lexical density.

```{r Load dependency}
library(data.table)
library(dplyr)
library(ggplot2)
library(RColorBrewer)
library(stringr)
library(stringi)
library(tm)
library(tidytext)
library(wordcloud2)
library(htmlwidgets)
```

```{r Load data}
chat_sessions <- fread("data/merge_files3.txt", sep = "\t")
```


## Clean data before tokenization

In this section, 1) contractions, 2) 1-3 character words, 3) stop words, and 4) undesirable words will be removed. Undesirable words include first names, "[Agent]Xx:", "[Visitor]Xx:", "comm100" and "has left the chat" and these will be removed.

```{r Prior cleaning}
# create list of first names that can be removed when filtering 
firstnames <- c(unique(unlist(stri_extract_all_regex(chat_sessions[, Chat], "\\[Agent\\][^\\s]+"))),
                unique(unlist(stri_extract_all_regex(chat_sessions[, Chat], "\\[Visitor\\][^:]+"))))

# remove NA
firstnames <- firstnames[!is.na(firstnames)]

# remove trailing spaces
firstnames <- str_trim(firstnames)

# remove Agent/Visitor
firstnames <- gsub("\\[Agent\\]|\\[Visitor\\]", "", firstnames)
firstnames <- gsub("^[[:punct:]]", "", firstnames)

# remove empty elements
firstnames <- firstnames[firstnames != ""]

# make lowercase
firstnames <- tolower(firstnames)

# make data.frame
firstnames <- data.frame(firstnames)
colnames(firstnames) <- "word"

# fix contractions
chat_sessions[, Chat := gsub("won't", "will not", Chat)]
chat_sessions[, Chat := gsub("can't", "can not", Chat)]
chat_sessions[, Chat := gsub("n't", " not", Chat)]
chat_sessions[, Chat := gsub("'ll", " will", Chat)]
chat_sessions[, Chat := gsub("'re", " are", Chat)]
chat_sessions[, Chat := gsub("'ve", " have", Chat)]
chat_sessions[, Chat := gsub("'m", " am", Chat)]
chat_sessions[, Chat := gsub("'d", " would", Chat)]
chat_sessions[, Chat := gsub("'s", "", Chat)]
chat_sessions[, Chat := gsub("s'", "s", Chat)]

# remove undesirable words
chat_sessions[, Chat := gsub("has left the chat", "", Chat)]

# remove agent and visitor and punctuation
chat_sessions[, Chat := gsub("\\[Agent\\][^:]+|\\[Visitor\\][^:]+|[[:punct:]]|[[:digit:]]", "", Chat)]
# chat_sessions[grep("[']", Chat)]

chat_sessions[, Chat := gsub("comm", "", Chat, ignore.case = T)]
  
# change to lowercase
chat_sessions[, Chat := tolower(Chat)]
```

## Tokenization using `tidytext` package

```{r Tokenize words}
# tokenize
chat_sessions_filtered <- chat_sessions %>%
  unnest_tokens(word, Chat) %>%
  anti_join(stop_words) %>%
  anti_join(firstnames) %>%
  distinct() %>%
  filter(nchar(word) > 3)

chat_sessions_filtered <- data.table(chat_sessions_filtered)

# what does the data frame looks like?
dim(chat_sessions_filtered)
```

There are `r dim(chat_sessions_filtered)[1]` words in total for `r chat_sessions_filtered[, length(unique(ID))]` chat sessions.  

## Explore chat sessions that did not pass the filter

```{r Chat session that did not pass filter}
fail_filter <- chat_sessions[!(ID %in% chat_sessions_filtered[, unique(ID)])]

fail_filter[, head(Chat)]
```

A quick check showed that these chat sessions are typically training sessions or the visitor has left the chat without replying. 

## Word frequency

In this section, the word frequency of each chat session is plotted. 

```{r Plot distribution of word frequency for chat sessions, fig.width=14, fig.height=8}
word_frequency <- chat_sessions_filtered[, .N, by = "ID"]

ggplot(word_frequency, aes(x=N)) +
    geom_histogram(binwidth=.5, colour=brewer.pal(11, "RdBu")[10], fill="white") +
  theme_bw() +
  theme(axis.text = element_text(size = 28),
        axis.title = element_text(size = 30, face = "bold")) +
  ylab("Count") +
  xlab("Word Frequency per Chat Session") +
  geom_vline(aes(xintercept=median(word_frequency[ ,N])),
             color="red", linetype="dashed", size=1) +
  scale_x_continuous(breaks = seq(0, 220, 20))
```

Median word frequency per chat session is `r word_frequency[, median(N)]`.

## Top words

### What are the top 20 words with the highest frequency? 

```{r Top 20 words}
top_words <- chat_sessions_filtered[, .N, by = "word"][order(-N)]

top_words[1:100]
```

```{r Plot top 20 words}
top20 <- top_words[1:20]

top20[, word := factor(word, levels = word)]

ggplot(data = top20, aes(x = word, y = N)) +
  geom_bar(stat="identity", position=position_dodge(), fill=brewer.pal(11, "RdBu")[10]) +
  coord_flip() +
  theme_bw() +
  theme(axis.text = element_text(size = 28),
        axis.title = element_text(size = 30, face = "bold"),
        axis.text.x = element_text(size = 24, angle = 45, hjust = 1)) +
  scale_y_continuous(breaks = seq(0, 1600, 200)) +
  scale_x_discrete(limits = rev(levels(top20[, word]))) +
  ylab("Frequency") +
  xlab("")
```

### Visualize top 300 words using wordcloud

```{r Plot top 300 words using wordcloud, fig.show='hold'}
wordcloud2(top_words[1:300], size = .5)
```

```{r Plot top 300 words, results='hide'}
wordcloud <- wordcloud2(top_words[1:300], size = .5)
results_dir <- paste0(getwd(), "/plots")
saveWidget(wordcloud, paste0(results_dir, "/wordcloud.html"), selfcontained = F)
```

